---
title: "chapter 4"
author: "Maika"
date: "13 helmikuuta 2017"
output: html_document
---

#Clustering and classification
##Classification

```{r}
library(MASS)
data("Boston")
str(Boston)
dim(Boston)

```

Here is our Boston dataset, which is one of R's own datasets. It conserns about Housing values in the suburbs of Boston, USA. It has 14 variables to explain the values and 506 observations, which consern about the town.



```{r}
summary(Boston)
```
The summary of our 14 variables show that all but 2 of our variables are numeric and continuous. We have binary variables chas, which gives 1 if tract is situated by the Charles river, and we also have rad, which gives and index for accesibility for radial highways.

*Visualizing correlation*


```{r}
library(dplyr); library(MASS); library(corrplot)
cor_matrix<-cor(Boston) %>% round(digits=2)
cor_matrix
corrplot(cor_matrix, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex=0.6)
```

Here we can see correlations between all the variables visualized as pairs. Only the upper side is needed, since it's symmetric. The bigger the dot and stronger the color, the stronger the correlation, too. 

Strong negative correlation seems to be between 1. medv and lstat, 2. age and dis, 3. nox and dis and 4. indus and dis. This means that 1: theless there is lower status inhabitants the higher value of the owner-occupied homes, 2: the younger the people the closer the Boston employments centres are, 3: and also, the closer to employmetn centres the less nitrogen oxides and lastly 4: the closer to employment centre the less industrial workers in the area.

The strongest positive correlation is between tax and rad (corr. 0.91). The better contacts with radial highways, the bigger property tax-rate.

Because these have the strongest correlations, I belive these are the ones we need to consentrate later on.

*Scaling the data*

We use scale to standardize our data. The variables communicate then better with eash other, so to speak. WE also want our new scales as data frame instead of matrix.


```{r}
boston_scaled <- scale(Boston)
summary(boston_scaled)
class(boston_scaled)
boston_scaled<- as.data.frame(boston_scaled)
boston_scaled
```
Because we have scaled our data, all observations are between integral 10. Now they can be handles together and seen the real sitances of these groups. This will help us in the clustering part.

But first we need to create new crime rate to our data. 

```{r}
boston_scaled <- scale(Boston)
boston_scaled<- as.data.frame(boston_scaled)
scaled_crim <- boston_scaled$crim
summary(scaled_crim)
crim_quant <- quantile(scaled_crim)
crime <- cut(scaled_crim, breaks = crim_quant, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))

table(crime)
boston_scaled <- dplyr::select(boston_scaled, -crim)

boston_scaled <- data.frame(boston_scaled, crime)
```

After our scaling we are able to make new variable crime to our dataset. WE use crim in our scaled boston dataset and use cut to adjust it to create categorical variable crime. WE have categories low, med low, med high and high based on quantiles of the crim. So there's 127 observations in the crimne low category, which are all the observations under 25% of all observations in crime. Since the old crim isn't necessary anymore, we can drop that out from our boston scaled.

Before we actually can use our model for clustering, we have to make sure that we can test how well our model works: That's why we're making train set to create our model and test set to predict our model to the test set. No I'll divide our dataset, so 80% will be used as train set and 20% to test set.

WE also want to save our crime data from the test dataset, so we can use it as our target variable.

```{r}
n <- nrow(boston_scaled)
random <- sample(n,  size = n * 0.8)
# train set
train <- boston_scaled[random,]
# test set 
test <- boston_scaled[-random,]
#save crime from test
correct_classes <- test$crime
# remove crime from test data
test <- dplyr::select(test, -crime)
```

Now we are ready to use Linear Discriminant Analysis to cluster our data. LDA is a classification method, where we need to know the classes before hand, but they can be trained with our data.

In our lda we have crime as target variable against all other variables. Here we obviously use our train set to create the model.

Then predict the classes with the LDA model on the test data. Cross tabulate the results with the crime categories from the test set. Comment on the results. (3 points)

```{r}
# linear discriminant analysis
lda.fit <- lda(formula= crime ~ ., data = train)

lda.fit

# the function for lda biplot arrows
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "red", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}

# target classes as numeric
classes <- as.numeric(train$crime)

# plot the lda results
plot(lda.fit, dimen = 2, col=classes, pch=classes)
lda.arrows(lda.fit, myscale = 1)
```

We want to find out what is the formula to predict crime rate in Boston. THe red arrows in the middle are the variables that are used to find crime rates. Now lets use our test set to see if we our model really predicted the crime rates correctly.



```{r}
# predict classes with test data
lda.pred <- predict(lda.fit, newdata = test)

# cross tabulate the results
table(correct = correct_classes, predicted = lda.pred$class)
```
Here is a cross tabulation with crime categories our model predicted against the real categories.

When the real value was low, the model predicted half of the time to low, but almost as many times to med_low and 2 instances to med_high. Almost same situation is also with med_low: half of the time our model is right and half is not. Then again med high is almost the same in correct observations as in our model and high crime rate is perfectly predicted by our model. So high end of crime rate is definitely better predicted than low end.

##Clustering

Now we scale again our data to use it for clustering method. Again, we need scaling to make distances measurable. 
*Scaling again*
```{r}
data("Boston")
boston_scaled2 <- scale(Boston)
summary(boston_scaled2)
class(boston_scaled2)
boston_scaled2<- as.data.frame(boston_scaled2)
boston_scaled2
```

*Distances calculated by euclidean method*
With clustering we do not know the number of proper classes before hand, so we need to find them out by using distance method. I am using the standard euclidean distance method here.

```{r}

dist_eu <- dist(boston_scaled2)

# look at the summary of the distances
summary(dist_eu)


```
*k-means*
k-means clusters our observations into certain number of clusers, that it calculates the most preferable. It clusters by grouping observations to the nearest mean. K-mean tries to find the optimal amount of clusters from the data. With function pairs we see a sharp drop when there´s 2 clusters. That seem to be the optimal number of clusters to our data according to k-means method.

```{r}
library(ggplot2); library(GGally)
set.seed(123)

# euclidean distance matrix

# determine the number of clusters
k_max <- 10

# calculate the total within sum of squares
twcss <- sapply(1:k_max, function(k){kmeans(dist_eu, k)$tot.withinss})

# visualize the results
plot(1:k_max, twcss, type='b')

# k-means clustering
km <-kmeans(dist_eu, centers = 2)

boston_scaled2$km <- as.factor(km$cluster)

# plot the Boston dataset with clusters
ggpairs(boston_scaled2, ggplot2::aes(colour=km))

```


##Conclusions

In this exercise we had two different methods to divide our data into groups based on their distance. In our first method (Classification), we divided the data by 4 classes of crime waves. It looked like it was quite accurate in the high end of crime but did not predict so well in low and mid low classes of criminality.

In the second method we used clustering based on no assumption of the classes and tried to use k-means to find those classes. K-means found two classes to be optimal to separate the data, which we can se in our pairs. The only problem is that we do not know without LDA what is the method to separate the two classes,so we do not have a good way of saying if the 2 classes given from k-means are meaningful according the data or not.


