---
title: "Chapter 3"
author: "Jaakko Kein??nen"
date: "11/21/2017"
output: html_document
---
#Chapter 3. RStudio Exercise 3
Read data to "alc" and check the column names.
```{r}
alc <- read.csv("/Users/caroram/Documents/Jaakon tutkimus/Open data kurssi s2017/GitHub/IODS-project/data/create_alc.csv")
colnames(alc)
str(alc)
```
Four variables that could be associated with alcohol use and hypotheses: 
1) sex: boys might drink more alcohol tha girls
2) age: those who are old enough might drink more
3) studytime: those who study more might dring less
4) famrel: those with good family relationships might drink more
Check distributions and correlations.
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
g1 <- ggplot(data = alc, aes(x = alc_use)) + geom_bar(aes(fill=sex), position="dodge")
g1
g2 <- ggplot(data = alc, aes(x = studytime)) + geom_bar()
g2
g3 <- ggplot(alc, aes(studytime, alc_use)) + geom_boxplot(aes(group=studytime))
g3

```
#logistic regression to predict alcohol high use with preselected 4 variables:
 
First change the studytime and famrel variables as factors (categorical).
```{r}
alc$studytime <- factor(alc$studytime)
alc$famrel <- factor(alc$famrel)
m <- glm(high_use ~ age + studytime + sex + famrel, data = alc, family = "binomial") 
summary(m)
```
Interpretation: Age is a significant predictor of high use (p<0.05). So is studytime category 3 (compared to category 1 which is not shown). This meand studying 5-10 hours weekly is associated with less high use than studying <2 hours. Category 4 (studying more than 10 hours weekly) was not significantly associated with high use. Male sex increased the probability of high use significantly. Family relations are not associated with high use.
#Odds ratios and confidence intervals:

```{r}
OR <- coef(m) %>% exp
CI<- confint(m) %>% exp
cbind(OR, CI)
```
CI:s of predictors do not include 0. 
```{r}
# predict() the probability of high_use
probabilities <- predict(m, type = "response")

# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability>0.5)

# see the last ten original classes, predicted probabilities, and class predictions
select(alc, studytime, famrel, sex, high_use, probability, prediction) %>% tail(10)

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction)
# initialize a plot of 'high_use' versus 'probability' in 'alc'
g <- ggplot(alc, aes(x = probability, y = high_use, col=prediction))

# define the geom as points and draw the plot
g + geom_point()

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table %>% addmargins()
# define a loss function (mean prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# call loss_func to compute the average number of wrong predictions in the (training) data
loss_func(class = alc$high_use, prob = alc$probability)
```

